---
title: "Classwork Biggish Data"
format: html
editor: visual
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r, include=FALSE}
library(tidyverse)
```

## Thanks to Dr. James Disckens.
### Instructions: Generate Rmarkdown file and a Word file that shows all required code and output.

```{r, message = FALSE}
library(tidyverse)    
library(dplyr)  
library(data.table)
```



  1. Use and show data.table code to select the variables year, month, day, and hour from the imported flights data.\
  - Load the file.\
  - Check its class.\ 
  - How many rows or columns it has?\
```{r, message = FALSE}
flights <- fread("nycdata.csv") 
```

```{r, eval=FALSE}
class(flights)
```

```{r, eval=FALSE}
glimpse(flights)
```
 
```{r, eval=FALSE}
names(flights)
```

```{r,eval=FALSE}
unique(flights$dest)
```

```{r,eval=FALSE}
unique(flights$origin)
```
- Select columns year, month, day, and hour 
 
```{r}
flights[,c("year", "month", "day", "hour")]
# Note there is no dot before c()
```

The (,) after [ means to use the entire rows. 
or

```{r,eval=FALSE}
flights[,.(year, month, day, hour)]
# There is a dot before (
```
or

```{r}
flights[,year:day]
```
- **For practice:  by tidyverse**
 
```{r}
# Read the file by tidyverse use read_scv()
df1 <- read_csv("nycdata.csv")
```

 - Note we may get the following message
```{block}
Use `spec()` to retrieve the full column specification for this data.
ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.
```
- There are three types of message.\
  - Errors: If the message is reporting an error we must fix it.\
  - Warnings: You can ignore it, but eventually you must fix it. You may search on line how to fix that particular warning.\
  - Notes: Minor problems that you may or may not fix it. But it is better to fix. It is better to fix.
 
```{r}
glimpse(df1)
```
 

```{r}
df1 %>%
  select(year,month,day, hour)
```
 
  

  2. Use and show data. table code to produce a table that shows a carrier of DL, an origin of JFK and a destination of SEA
  
```{r}
flights[c(origin == "JFK"& carrier == "DL"& dest == "SEA")]
```


```{r}
flights[origin == "JFK"& carrier == "DL"& dest == "SEA"]
```

  3. Use and show data.table code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
  
```{r}
unique(flights$month)
```
  
  
```{r}
flights[carrier== "UA" & month == 3 & air_time < 330]
```
 
```{r}
flights[carrier== "UA" & month == 3 & air_time < 330, .(carrier, month, air_time)]
```
```{r}
flights2 %>%
  filter(carrier== "UA" , month == 3 & air_time < 330) %>%
  select(carrier, month , air_time)
```
  

  4. Use and show tidyverse code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
```{r}
flights1 <- read_csv("nycdata.csv") 
glimpse(flights1)
```

```{r}
class(flights1)
```


```{r}
flights1 %>%
  filter(carrier== "UA" & month == 3 & air_time < 330)
```
  
```{r}
flights1 %>%
  filter(carrier== "UA" & month == 3 & air_time < 330) %>%
  select(carrier, month, air_time)
```
  

  5. Use the data.table method to add a variable called speed that is the average air speed of the plane in miles per hour. 

```{r}
names(flights)
```

```{r}
glimpse(flights)
```

  
```{r}
flights[,c("speed") := (distance/hour)]
flights
```

```{r}
flights[,.(speed, distance, hour)]
```


or

```{r}
flights[,c("Speed") := (60*distance/air_time)]
flights
```




```{r}
flights[,.(Speed, distance, air_time)]
```

- Or by chaining

```{r}
flights[,c("Speed") := (60*distance/air_time)][,.(Speed, distance, air_time)]
```


  6. Use the tidyverse method to add a variable called speed that is the average air speed of the plane in miles per hour.
  
```{r}
flights1 %>%
  mutate(speed = (distance/hour))
```

```{r}
flights1 %>%
  mutate(speed = (distance/hour)) %>%
  select(speed, distance, hour)
```
  7. Show and use coding to change the carrier abbreviation of UA to UniitedAir,\
    7a. data.table method\


```{r}
unique(flights$carrier)
```
```{r}
flights[carrier== "UA", carrier:= "UniitedAir" ]
```

```{r}
unique(flights$carrier)
```
    7b. tidyverse method  (Use a sequence of dplyr commands so that you can see the change in your table)
    
```{r}
unique(flights1$carrier)
```

```{r}
flights1 %>%
  mutate(carrier= recode(carrier,"UA" = "UniitedAir")) ->
  flights1
flights1
```

```{r}
unique(flights1$carrier)
```

