---
title: "Basic Stat Lab"
format: html
editor: visual
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3.5, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

# Basic Statistics Classwork/Lab

1)  In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a -- e) You can use any method for part f.\

    a)  What proportion of the Hong Kong population is between 170 cm and 180 cm?\

    ```{r}
    high <- pnorm(180, 171.5, 5.5) 
    low <- pnorm(170, 171.5, 5.5)
    high - low
    ```

    b)  What proportion of the Hong Kong population is less that 165 cm ?\

    ```{r}
    pnorm(165, 171.5, 5.5)
    ```

    c)  What proportion of the Hong Kong population is greater than 175 ?\

    ```{r}
    pnorm(175, 171.5, 5.5, FALSE)
    ```

    d)  The proportion .25 is less than what male height value ?\

    ```{r}
    qnorm(0.25, 171.5, 5.5)
    ```

    e)  The proportion .85 is greater than what male height value ?\

    ```{r}
    qnorm(0.85, 171.5, 5.5, FALSE)
    ```

    f)  The proportion of .68 is between what two male height values ?\

    ```{r}
    low <- qnorm(.16, 171.5, 5.5)
    high <- qnorm(.84, 171.5, 5.5)
    low
    high

    # checking that it equals 0.68
    pnorm(high, 171.5, 5.5) - pnorm(low, 171.5, 5.5)
    ```

2)  Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.

```{r}
dnorm(2.5)
```

3)  Use and show R code to determine the height of the standard normal curve at a Z value of .4.

```{r}
dnorm(.4)
```

4)  Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.

```{r}
dt(1.5, 3)
```

5)  Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.

```{r}
dt(-2, 2)
```

6)  For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.

```{r}
pt(2.5, 3)
```

7)  For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.

```{r}
pt(1.75, 1, lower.tail = FALSE)
```

8)  For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.

```{r}
qt(0.355, 2, lower.tail = FALSE)
```

9)  An educator believes that new directed reading activities in the classroom will help elementary school pupils improve some aspects of their ability. She arranges for a third -grade class of 23 students to take part in these activities for an eight-week period. A control classroom of 23 third graders follows the same curriculum without the activities. At the end of the eight weeks, all students are given a Degree of Reading Power (DRP) test, which measures the aspects of reading ability that the treatment is designed to improve. The sample data performance results are provided below;

-   Treatment Group: 24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70

-   Control Group: 42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42

-   Design and execute a two sample t test.\

    a)  State the appropriate null and alternative hypotheses.

    ```{r}
    print("The null hypothesis is that the treatment and control groups will have the same test scores. The alternative hypothesis is that the treatment group scores higher than the control group.")
    ```

    b)  Use and show R code to produce the p value and the confidence interval.

    ```{r}
    Treatment <- c(24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70)
    Control <- c(42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42)

    library(broom)
    library(tidyverse)
    t.test(Treatment, Control, "greater") %>% 
      tidy() -> result

    print(paste0("The p-value is: ", result$p.value, ". The confidence interval is: ", result$conf.low, " - ", result$conf.high, ".", sep = ""))
    ```

    c)  Determine if you should reject or fail to reject the null hypothesis using the p value and the confidence interval.

    ```{r}
    print("Because the p-value is 0.0053, I will reject the null hypothesis. The confidence interval states the pupils with new directed reading activities will score atleast 4.38 points better on the test than if they did not.")
    ```
